---
title: "Creating a Demo Data frame"
author: "geoKelpMod"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
editor_options: 
  markdown: 
    wrap: 72
---

# Steps to create a demo raster data frame for kelp calculation

Steps:

1.  Load the libraries

2.  Create the variables and values

3.  Merge and create the 2D Data Frame with a date column

4.  Print the data in time series ggplots

5.  Convert data frame into a raster (Kelp_Biomass)

6.  Create the polygon for the area of interest

7.  Create a tmap that shows the values.

### 1. Loading the libraries

```{r}
library(tidyverse)
library(raster)
library(sp)
library(tmap)
library(sf)
library(stars)
library(gstat)
library(rgdal)
library(kml)

```

### 2. Creating the random variables/ columns

```{r}
set.seed(35) # setting seed for reproducibility with our team number

latitude <- runif(100, min = 33.83, max = 34.49)
longitude <- runif(100, min = -120.50, max = -119.45)
kelp_biomass <- rnorm(100, mean = seq(300, 80, length.out = 100), sd = seq(50, 5, length.out = 100)) #adding trend with a sd from a sequence.
depth <- runif(100, min = 17, max = 40)
sst <- rnorm(100, mean = seq(10, 18, length.out = 100), sd = seq(2, 5, length.out = 100))
NO3 <- rnorm(100, mean = seq(0, 1, length.out = 100), sd = seq(0, 0.1, length.out = 100))
NO2 <- runif(100, min = 0.001, max = 0.05)
NH4 <- runif(100, min = 1, max = 5)
year <- rep(2014:2022, each = 12)[1:100]
quarter <- sample(1:4, 100, replace = TRUE)

```

### 3. Merge all the variables into a data frame and add a `date` column

```{r}
kelp_df <- data.frame(latitude, longitude, kelp_biomass, depth, sst, NO3, NO2, NH4, year, quarter)

# Create a new date column merging year and quarter columns
kelp_df$date <- as.Date(paste0(kelp_df$year, "-", as.character((kelp_df$quarter-1)*3+1), "-01"))

```

### 4. Print the variables as time series graphs

a)  In the same plot

```{r}
colors <- c("darkgreen", "#56B4E9", "orange", "purple", "#0072B2", "#D55E00")
y_vars <- c("kelp_biomass", "depth", "sst", "NO3", "NO2", "NH4")
# Create a plot for each variable with a buffer of 10% of the data range
plot_list <- lapply(seq_along(colors), function(i) {
  var <- y_vars[i]
  y_range <- range(kelp_df[[var]])
  buffer <- diff(y_range) * 0.2
  ggplot(kelp_df, aes(x = date, y = !!sym(var))) +
    geom_line(color = colors[i]) +
    labs(x = "Date", y = var) +
    ylim(y_range + c(-buffer, buffer))
})

# Combine the plots into a single figure using facet_wrap
do.call(gridExtra::grid.arrange, c(plot_list, ncol = 2))

```

b)  Print them in six different graphs

```{r}
# Create a vector of y variables to loop through
colors <- c("darkgreen", "#56B4E9", "orange", "purple", "#0072B2", "#D55E00")

# Create a for loop to generate a separate ggplot for each y variable
for (i in seq_along(y_vars)) {
  
  # Create the ggplot
  p <- ggplot(kelp_df, aes(x = date, y = !!sym(y_vars[i]))) +
    geom_line(color = colors[i]) +
    labs(title = y_vars[i], y = y_vars[i], x = "Date")
  
  # Print the ggplot
  print(p)
  
}
```

c)  Example of the Kelp Biomass graph....

```{r}
ggplot(kelp_df, aes(x = date, y = kelp_biomass)) +
  geom_line(color = "darkolivegreen") + 
  labs(title = "Kelp Biomass Time Series",
       subtitle = "Santa Barbara Channel",
       x = "Date",
       y = "Kelp Biomass (kg)",
       caption = "LTER")
```

### 5. Convert the data frame into a raster.

```{r}
# Convert kelp_df to sf object
kelp_sf <- st_as_sf(kelp_df, coords = c("longitude", "latitude"), crs = st_crs(4326))

# Create a raster template using the world map
r <- raster(extent(kelp_sf), nrow = 100, ncol = 100, crs = st_crs(4326))


# Rasterize the kelp_sf object
kelp_raster <- rasterize(kelp_sf, r, field = "kelp_biomass")
```


```{r}

st_raster_depth <- st_rasterize(kelp_sf |> select(depth))

```

```{r}
summary(kelp_raster)
plot(kelp_raster)
values(kelp_raster)
```


### 6. Create the polygon that shows the area we are interested on

```{r}
# Define the rectangle as a polygon
rectangle <- st_polygon(list(rbind(c(-120.50, 34.49), c(-120.50, 33.83), c(-119.45, 33.83), c(-119.45, 34.49), c(-120.50, 34.49))))

# Convert the polygon to an sf object
area_sf <- st_sf(geometry = st_sfc(rectangle))

# Set the CRS of the rectangle to match the kelp_raster
st_crs(area_sf) <- st_crs(kelp_raster)

```


### 8. Interpolation using the IDW method (inverse distance weighting)
http://132.72.155.230:3838/r/spatial-interpolation-of-point-data.html

```{r}

# convert to data.frame with x, y, and z columns
kelp_sf_2 <- st_as_sf(kelp_df, coords = c("longitude", "latitude"), crs = st_crs(4326))

# define the bounding box
xmin <- -120.50
ymin <- 33.83
xmax <- -119.45
ymax <- 34.49
crs <- st_crs("+proj=longlat")

# create a data.frame with the coordinates
coords <- data.frame(x = c(xmin, xmax), y = c(ymin, ymax))

# convert to an sf object and create a regular grid of points
grid_sf <- st_as_sf(coords, coords = c("x", "y"), crs = crs) %>%
  st_make_grid(cellsize = 0.03)  # adjust the cell size as needed

# interpolate using IDW method
idw_model <- gstat::idw(kelp_biomass ~ 1, kelp_sf_2, newdata = grid_sf, idp = 2)

# Create an empty raster with the desired resolution and extent
empty_r <- raster(extent(area_sf), resolution = res(kelp_raster))

# rasterize the idw_model object using the var1.pred column
idw_raster <- rasterize(idw_model, empty_r, field = "var1.pred")

# crop the raster to the area of interest
kelp_raster_cropped <- crop(idw_raster, extent(area_sf))

```


Create a for loop that rasterize each attribute from the kelp_df

```{r}
y_vars <- c("kelp_biomass", "depth", "sst", "NO3", "NO2", "NH4")

for (i in y_vars){
  
  # interpolate with IDW method per attribute
  attribute_model <- gstat::idw(depth ~ 1, kelp_sf_2, newdata = grid_sf, idp = 2)
  
  
  # rasterize the idw_model_object object using the var1.pred column
  attribute_raster <- rasterize(attribute_model, empty_r, field = "var1.pred")
  
  
  # crop the atteribute_raster to the area of interest
  cropped_raster <- crop(attribute_raster, extent(area_sf))
  
  # assign the cropped raster to a new variable with the unique name in the global environment
  return(paste0(i, "cropped_raster"))
  
}

```




### 9. Create a new model grid
```{r}

# Set the dimensions of the grid
nrow <- 100
ncol <- 100

# Create a grid with random values between 0 and 100
grid <- matrix(runif(nrow*ncol, min = 0, max = 100), nrow = nrow, ncol = ncol)

# Add bias towards 0
grid[grid < 96] <- rnorm(sum(grid < 80), mean = 0, sd = 5)

# Convert the grid to a raster object
model_raster <- raster(grid)

# Assign the extent of 'kelp_raster' to 'r'
extent(model_raster) <- extent(kelp_raster)

plot(model_raster)

```


### 10. Plot the models and data raster with tmap

```{r}
# Load the shapefile
channel_islands <- st_read(here::here("data", "channel_islands.shp"))

```


```{r}
# Plot world map, add the area of interest and the raster values
data("World")

tm_shape(World) +
  tm_polygons() +
  
tm_shape(channel_islands) +
  tm_polygons()+
  
tm_shape(area_sf) +
  tm_borders(lwd = 1, col = "red") +
  
tm_shape(kelp_raster_cropped) +
  tm_raster(style = "fisher", palette = "Greens", alpha = 0.8) +
  
tm_shape(model_raster) +
  tm_raster(style = "fisher", palette = "Blues", alpha = 0.8) +
  
  tm_legend(title = "kelpGeoMod") +
  tm_scale_bar(position = c("left", "bottom"), text.size = 0.7) +
  tm_view(set.view = c(-120, 34.1, 8))

tmap_mode("view")
```





```{r}

tm_shape(World) +
  tm_polygons() +
tm_shape(channel_islands) +
  tm_polygons()+
tm_shape(area_sf) +
  tm_borders(lwd = 1, col = "red") +
tm_shape(raster_kelpbiomass) +
  tm_raster() +
tm_shape(kelp_raster_cropped) +
  tm_raster() +
  tm_shape(idw_model) +
  tm_raster() +
  tm_view(set.view = c(-120, 34.1, 8))

tmap_mode("view")

```


